---
title: "Market Basket Analysis"
author: "Sean Wong"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```


```{r packages, include = F}
rm(list=ls())
gc()
setwd("")
require(tidyverse)
require(readxl)
require(ggplot2)
require(lubridate)
require(dplyr)
require(stringr)
require(openxlsx)
require(priceR)
# require(apriori)
```

```{r data_load, include = F}
currencies <- c("SGD", "MYR", "TWD", "AUD", "NZD", 
                "USD", "IDR", "JPY", "HKD")
df <- read.csv("Transactions Buyer.csv") %>% 
  filter(str_detect(Buyer, "@email.com") != T) %>%
  mutate(Created.At = as.Date(gsub("T.*", "", Created.At), "%Y-%m-%d"),
         Quarter = floor_date(Created.At, unit = "quarter"),
         Month = floor_date(Created.At, unit = "month")) %>% filter(Buy.Sell == "buy") %>% 
  filter(Buyer.Currency %in% currencies)

excl_status <- c("buy_failed", "canceled", "failed", 
                 "payment_failed", "sell_failed")
excl_pay_method <- c("credit")

df <- df %>% filter(Payment.Status == "authorized" | Payment.Status == "captured") %>% filter(!Status %in% excl_status) 

curr_start_date <- min(df$Created.At)
curr_end_date <- max(df$Created.At)

avg_exch <- NULL
name_avg_exch <- NULL

curr_start_date <- min(df$Created.At)
curr_end_date <- max(df$Created.At)

for (c in 1:length(currencies)){
  exch <- historical_exchange_rates(from = currencies[c], to = "SGD",
                                    start_date = curr_start_date,
                                    end_date = curr_end_date) %>% data.frame()
  current_curr <- currencies[c]
  exchange_rate <- mean(exch[,2], na.rm=T)
  
  avg_exch <- append(avg_exch, exchange_rate)
  name_avg_exch <- append(name_avg_exch, current_curr)
}

names(avg_exch) <- name_avg_exch

df <- df %>% mutate(
  Total.paid.local = ifelse(is.na(Total.paid.local) == T, Price.Paid..SGD., Total.paid.local),
  
  Price.Paid..SGD. = case_when(
    Buyer.Currency == "SGD" ~ Total.paid.local*avg_exch["SGD"],
    Buyer.Currency == "MYR" ~ Total.paid.local*avg_exch["MYR"],
    Buyer.Currency == "TWD" ~ Total.paid.local*avg_exch["TWD"],
    Buyer.Currency == "AUD" ~ Total.paid.local*avg_exch["AUD"],
    Buyer.Currency == "NZD" ~ Total.paid.local*avg_exch["NZD"],
    Buyer.Currency == "USD" ~ Total.paid.local*avg_exch["USD"],
    Buyer.Currency == "IDR" ~ Total.paid.local*avg_exch["IDR"],
    Buyer.Currency == "JPY" ~ Total.paid.local*avg_exch["JPY"],
    Buyer.Currency == "HKD" ~ Total.paid.local*avg_exch["HKD"],
    is.na(Buyer.Currency) == T ~ Total.paid.local)
)
```

Buyer Data is from ForestAdmin, from 1 Apr 2022 onwards.  
We are looking at `r currencies` and have excluded transactions with `r excl_status` statuses as well as payment methods which include `r excl_pay_method`.  

All buyers with @email.com have also been excluded. 
We are looking at Product ID instead of MPN/SKU as there are certain records with blank MPN/SKUs.  


# Start
```{r data_prep, include = F}
# SELECT RELEVANT COLUMNS HERE
df <- df %>% select(
  #demographics
  Buyer, Status, Buyer.Currency, 
  #amount
  Trxn.ID, Price.Paid..SGD., Product, Size, Product.ID,
  #date
  Created.At, Month, Quarter
) 

cleaning_countries <- function(dataframe){
  test <- dataframe
  country <- test %>% select(Buyer.Currency) %>% unique() %>% pull()
  
  test <- test %>% 
    group_by(Quarter, Buyer, Product.ID) %>% 
    summarise(trxn = n_distinct(Trxn.ID), 
              trxn = sum(trxn, na.rm=T), 
              gmv = sum(Price.Paid..SGD., na.rm=T)) %>% 
    arrange(Buyer) %>% mutate(Buyer.Currency = country)
  
  return(test)
}

countries <- df %>% group_by(Buyer.Currency) %>% group_split() %>%
  map_dfr(~cleaning_countries(.x))
  
```

```{r exploration, include = F}
# finding the average basket size of each country

basket_size <- function(dataframe){
  test <- dataframe %>% filter(complete.cases(.))
  country <- test %>% select(Buyer.Currency) %>% pull() %>% unique()
  
  basket <- test %>% group_by(Buyer) %>% #invoice no.
    summarise(n_total = n(), n_items = n_distinct(Product.ID)) %>% #Product = StockCode 
    # summarise(avg_total_items = mean(n_total), 
    #           avg_dist_items = mean(n_items)) %>% 
    mutate(Buyer.Currency = country)
  return(
    basket
  )
}

ggplot(countries %>% group_by(Buyer.Currency) %>% 
  group_split() %>% map_dfr(~basket_size(.x)) %>%
    filter(n_items < 20), aes(n_items)) +
  geom_bar() + ggtitle("Distribution of basket sizes") + facet_wrap(~Buyer.Currency, scales = "free")
```

We see that most countries have a basket size of . 

```{r, summary_basket_size}
basket_size_summ <- function(dataframe){
  test <- dataframe %>% filter(complete.cases(.))
  country <- test %>% select(Buyer.Currency) %>% pull() %>% unique()
  
  basket <- test %>% group_by(Buyer) %>% #invoice no.
    summarise(n_total = n(), n_items = n_distinct(Product.ID)) %>% #Product = StockCode 
     summarise(avg_total_items = mean(n_total), 
               avg_dist_items = mean(n_items)) %>% 
    mutate(Buyer.Currency = country)
  
  return(
    basket
  )
}
DT::datatable(
  countries %>% group_by(Buyer.Currency) %>% 
    group_split() %>% map_dfr(~basket_size_summ(.x)) %>% 
    select(Buyer.Currency, avg_total_items, avg_dist_items), 
  filter = c("top")
)
```


# High Rollers  
```{r, include = F}
high_rollers <- read.csv("RFyA_Full Data_data.csv") %>% 
  rename("Buyer" = "email") %>% select(Buyer) %>% unique() %>% pull()

high_roll_df <- df %>% filter(Buyer %in% high_rollers)
```

## Where are they from?  

```{r, include = F}
ggplot(high_roll_df %>% group_by(Buyer.Currency) %>% 
  summarise(txn = n_distinct(Trxn.ID, na.rm=T)), 
  aes(x=reorder(Buyer.Currency, -txn), y = txn, fill = Buyer.Currency, col = Buyer.Currency)) + geom_bar(stat="identity") + 
  labs(title = "Where are High Rollers from?", 
       subtitle = "From 1 Apr 2022 onwards")
```


## Purchasing Trend of High Rollers  
```{r, include = F}
coeff <- 1000
ggplot(high_roll_df %>% group_by(Month) %>% 
         summarise(txn = n_distinct(Trxn.ID, na.rm=T),
                   gmv = sum(Price.Paid..SGD. ,na.rm=T)), 
       aes(x=Month, y = gmv, col = "blue", fill = "blue")) + geom_bar(position = "dodge", stat = "identity") + 
  geom_line(aes(y = txn*coeff, col = "red", group = 1, linewidth = 1.2)) + scale_y_continuous(name = "GMV",                                                                         sec.axis = sec_axis(~./coeff, name = "txns")) + theme(legend.position = "blank") + labs(title = "Trxn & GMV Trend of High Rollers", subtitle = "Absolute; from 1 Apr '22")
```


```{r high_roll_trend, include = F}
ggplot(df %>% mutate(high_roll = ifelse(Buyer %in% high_rollers, "high_roll", "others")) %>% 
  group_by(Month, high_roll) %>% 
  summarise(txn = n_distinct(Trxn.ID, na.rm=T),
            gmv = sum(Price.Paid..SGD., na.rm=T)),
  aes(x=Month, y = txn, fill = high_roll, col = high_roll)) + 
  geom_bar(stat="identity", position = "fill") +
  labs(title = "% of Trxns")

ggplot(df %>% mutate(high_roll = ifelse(Buyer %in% high_rollers, "high_roll", "others")) %>% 
  group_by(Month, high_roll) %>% 
  summarise(txn = n_distinct(Trxn.ID, na.rm=T),
            gmv = sum(Price.Paid..SGD., na.rm=T)),
  aes(x=Month, y = gmv, fill = high_roll, col = high_roll)) + 
  geom_bar(stat="identity", position = "fill") +
  labs(title = "% of GMV")
```

```{r, include = F}
# what do they normally buy? 
DT::datatable(high_roll_df %>% group_by(Product, Product.ID) %>% 
  summarise(txn = n_distinct(Trxn.ID, na.rm=T),
            gmv = sum(Price.Paid..SGD., na.rm=T)) %>% 
  arrange(desc(txn)))
```

# MBA  
These are the top 10 items that have the `highest support`.  
+ *Support:* The portion of our item set in our dataset occurs. Support tells us what percentage of transactions contain the combination of items A and B. It assists in identifying combinations that are frequent enough to be of interest (e.g., purchasing fish alone or purchasing fish and lemons together).  

+ *Confidence:* the probability that a rule is correct with items on the left for a new transaction. Confidence tells us what percentage of transactions with item A also have item B. (e.g., how many transactions that have bread also have butter).  

+ *Lift:* The ratio that exceeds the expected confidence by the confidence of the rule. The ratio of the number of respondents obtained with the model to the number obtained without the model is known as lift.  

*Lift (A => B) = 1* means that within a set of elements there is no correlation.  

*Lift (A = > B) > 1* means a positive correlation is more commonly purchased between the products in the product set, i.e. in items A and B.  

*Lift(A =>B) <1* means that it is unlikely to be purchased together for the negative correlation of the itemset, i.e. the products in the item set, A, and B.  

The top few items:  

```{r rules}
require(arules)
require(arulesViz)
df_clean <- df %>% group_by(Buyer, Product.ID) %>% 
  summarise(txn = n_distinct(Trxn.ID, na.rm=T)) 


countries_list <- split(df_clean$Product.ID,
                        df_clean$Buyer) 
online_trxn <- as(countries_list, "transactions")
print(summary(online_trxn))
```
```{r, include = F}
itemFrequencyPlot(online_trxn, topN=10, type = "absolute")
```


These are the top 10 items that have the `highest support`.  

```{r freqitems}
itemFrequencyPlot(online_trxn, topN=10, type = "absolute")
```

At a 10% confidence, we get the following:  

```{r rules1}
# the minimum support is 0.001, and the minimum confidence is 0.5 
rules <- apriori(online_trxn, parameter = list(supp = 0.001, conf = 0.1))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```

At a 5% confidence, we get the following:  

```{r rules0.05}

rules <- apriori(online_trxn, parameter = list(supp = 0.001, conf = 0.05))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```
  
  
## Running without Product X, Y, Z (Skews the data)  

At a 5% confidence, we get the following:  
```{r freq_nopanda}

df_noprod <- df_clean %>% 
  filter(!Product.ID %in% c(
    47163, #product X
    47054, #product Y
    47298)) #product Z

countries_list_noprod <- split(df_noprod$Product.ID, df_noprod$Buyer)
online_trxn_noprod <- as(countries_list_noprod, "transactions")

rules <- apriori(online_trxn_noprod, parameter = list(supp = 0.001, conf = 0.05))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```  

At a 1% confidence level:  

```{r}
rules <- apriori(online_trxn_noprod, parameter = list(supp = 0.001, conf = 0.01))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```

## What if we look at a Quarterly basis? 

At 10%:  

```{r}
df_qtr <- df %>% group_by(
  Buyer, Product.ID, Quarter) %>%
  summarise(txn = n_distinct(Trxn.ID, na.rm=T))

qtr_countries_list <- split(df_qtr$Product.ID,
                        df_qtr$Buyer) 
qtr_online_trxn <- as(qtr_countries_list, "transactions")

rules <- apriori(qtr_online_trxn, parameter = list(supp = 0.001, conf = 0.1))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```

At 5%:  
```{r}
rules <- apriori(qtr_online_trxn, parameter = list(supp = 0.001, conf = 0.05))
rules <- sort(rules, by = "lift", decreasing = T)

options(digits = 3)
print(inspect(rules))
```

```{r}
plot(rules, method = "graph")
```